# Differentially Expressed Genes

These are visualizations of the differentially expressed genes (DEGs).

## Top DEGs

These are the top 50 significantly (adjusted p-value <= 0.05) differentially expressed genes (DEGs) for each comparison determined by greatest absolute log2FoldChange.

```{r degs-topdegs}
tryCatch({
# Pull list of named DEG tables from global state
global_state$data %>%
  purrr::map("data") %>%
  combined_deg_heatmap(df = .)
}, error = function(e) {
    # This code runs only if an error occurs
    message("An error occurred, returning NULL.")
    return(NULL)
  })

```

## Volcano Plots

These are volcano plots of the DEGs.

```{r degs-volcanoplots, results="asis"}
# Dynamically generate volcano plots for each data set
# Load the configuration (usually at the start of your index.Rmd or setup chunk)
config <- yaml::read_yaml("configuration.yml")

# Access the toggle
show_fdr <- config$use_fdr

global_state$data %>%
  purrr::map(~ {
    # Create a new environment for each child
    child_env <- new.env(parent = environment()) 
    child_env$current_data <- .$data 
    child_env$current_name <- .$name 

    # Determine if data is valid for plotting
    child_env$has_data <- !is.null(.$data) && is.data.frame(.$data) && nrow(.$data) > 0

    knitr::knit_child(
      text = c(
        '',
        '### `r current_name`', 
        '',
        '```{r, echo=FALSE}',
        'if (has_data) {',
        '  volcano_plot(current_data, show_fdr)',
        '} else {',
        '  cat("No significant data available to display for this comparison.")',
        '}',
        '```',
        ''
      ),
      envir = child_env,
      quiet = TRUE 
    )
  }) %>%
  purrr::flatten_chr() %>%
  cat(sep = "\n")
```

<!-- ## Expression of Top DEGs -->
```{r degs-degexpression, results='asis'}
# Dynamically generate heatmaps for each data set, if raw counts are included in the analysis

if(global_state$using_counts) {
  global_state$data %>%
    map( ~ knitr::knit_child(
      text = c(
        '',
        '### `r .$name`',
        '',
        '```{r}',
        'group_count_heatmap(.)',
        '```',
        ''
      ),
      envir = environment(),
      quiet = TRUE
    )) %>%
    c("\n\nThis is gene expression of the most upregulated (top 90%) and most downregulated (bottom 10%) significant genes (FDR adjusted p-value < 0.05) \n\n", .) %>%
    c("\n\n## Expression of Top DEGs\n\n", .) %>%
    purrr::list_c() %>%
    cat(sep = "\n")
}
```
